---
title: "Hybrid health regimes: Access to primary care physicians and COVID-19 vaccine uptake across municipalities in Romania"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---
# Setup
Import libraries. 

```{r}
library(readxl) # read excel files
library(jtools) #for robust errors
library(sjPlot) #for tab_model
# Model comparison
library(performance) # model performance
library(lmtest) #Log likelihood ratio test
library (olsrr) #Collinearity Detection
# Spatial Regression 
library (sp) # dependency for rgdal
library(spdep) # for spatial weights generation
library (spatialreg)  # spatial regression
library (sphet) # spatial lag and error HET
library (car) #for VIF
```
Set the number of digits to 3, and scientific notation to 7
```{r}
options(digits=3, scipen=3)
```
Read the dataset (non-spatial one):
```{r}
ro_Vaccines_incidence_socioec <- read.csv(file = "data/ro_Vaccines_incidence_socioec_extended.csv")
head(ro_Vaccines_incidence_socioec)
```
Read the dataset (spatial one, spatial weights matrix):
```{r}
w_queen <- read.gal("data/queen_1ord_symmetric.gal")
w_queen <- nb2listw(w_queen, style='W')
summary(w_queen)
```
Rename columns:
```{r}
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[2]] <- 'MaxCovid'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[5]] <- 'VaccinationRateT1'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[8]] <- 'VaccinationRateT2'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[48]] <- 'FamilyDoctors'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[17]] <- 'Employees'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[53]] <- 'ElderlyRetired'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[40]] <- 'Migration'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[51]] <- 'SocialistInv'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[25]] <- 'RelativePoverty'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[52]] <- 'ElderlyMinimum'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[50]] <- 'Roma'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[31]] <- 'LivingSpace2020'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[23]] <- 'PrimaryEd4'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[39]] <- 'Neoprotestants'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[27]] <- 'VotesYES2018'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[43]] <- 'Urban'
names(ro_Vaccines_incidence_socioec)[names(ro_Vaccines_incidence_socioec) == colnames(ro_Vaccines_incidence_socioec)[47]] <- 'Size100'
```
Rename dataset:
```{r}
spatdata <- ro_Vaccines_incidence_socioec
head(spatdata)
```
# Non-spatial modelling
Formula declarations:
```{r}
eq1 = VaccinationRateT1 ~ FamilyDoctors + Employees + ElderlyRetired + Migration + Size100 + SocialistInv + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020 + Neoprotestants + VotesYES2018 + MaxCovid + Urban
eq2 = VaccinationRateT1 ~ FamilyDoctors +                            + Migration + Size100 + SocialistInv + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020 + Neoprotestants + VotesYES2018 + MaxCovid + Urban
eq3 = VaccinationRateT2 ~ FamilyDoctors + Employees + ElderlyRetired + Migration + Size100 + SocialistInv + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020 + Neoprotestants + VotesYES2018 + MaxCovid + Urban
eq4 = VaccinationRateT2 ~ FamilyDoctors +                            + Migration + Size100 + SocialistInv + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020 + Neoprotestants + VotesYES2018 + MaxCovid + Urban
```

```{r}
olsT1A <- lm(eq1, data=spatdata)
olsT1B <- lm(eq2, data=spatdata)
olsT2A <- lm(eq3, data=spatdata)
olsT2B <- lm(eq2, data=spatdata)

summ (olsT1A, digits=3, robust=FALSE, vifs=TRUE)
summ (olsT1B, digits=3, robust=FALSE, vifs=TRUE)
summ (olsT2A, digits=3, robust=FALSE, vifs=TRUE)
summ (olsT2B, digits=3, robust=FALSE, vifs=TRUE)
```
Fit measures
```{r}
compare_performance (olsT1A, olsT1B, olsT2A, olsT2B)
logLik(olsT1A)
logLik(olsT1B)
logLik(olsT2A)
logLik(olsT2B)
lrtest (olsT1A, olsT1B, olsT2A, olsT2B) 
```
Breusch Pagan Test for Heteroskedasticity
```{r}
ols_test_breusch_pagan(olsT1A)
ols_test_breusch_pagan(olsT1B)
ols_test_breusch_pagan(olsT2A)
ols_test_breusch_pagan(olsT2B)
```
Multicollinearity Condition Number
```{r}
ols_eigen_cindex(olsT1A)
ols_eigen_cindex(olsT1B)
ols_eigen_cindex(olsT2A)
ols_eigen_cindex(olsT2B)
```
Multicollinearity VIF - Variance Influence Factor
```{r}
car::vif(olsT1A)
car::vif(olsT1B)
car::vif(olsT2A)
car::vif(olsT2B)
```
Moran test for olsT1 & olsT2
```{r}
lm.morantest(olsT1A,w_queen)
lm.morantest(olsT1B,w_queen)
lm.morantest(olsT2A,w_queen)
lm.morantest(olsT2B,w_queen)
```
Lagrange multiplier olsT1 & olsT2
```{r}
lm.LMtests(olsT1A,w_queen,test="all")
lm.LMtests(olsT1B,w_queen,test="all")
lm.LMtests(olsT2A,w_queen,test="all")
lm.LMtests(olsT2B,w_queen,test="all")
```
# Spatial modelling
Model estimation for SARAR-HET Durbin
Model spatial lag specification search via simulations
```{r}
lagged_x_list <- c(Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  RelativePoverty +  ElderlyMinimum +  Roma,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  RelativePoverty +  ElderlyMinimum +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  RelativePoverty +  PrimaryEd4 +  Roma,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  RelativePoverty +  PrimaryEd4 +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  RelativePoverty +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  ElderlyMinimum +  PrimaryEd4 +  Roma,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  ElderlyMinimum +  PrimaryEd4 +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  ElderlyMinimum +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  Migration +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  Roma,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  RelativePoverty +  ElderlyMinimum +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  RelativePoverty +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  ElderlyRetired +  ElderlyMinimum +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  Migration +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  Roma,
                   Durbeq =  ~ FamilyDoctors +  Employees +  Migration +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  Migration +  RelativePoverty +  ElderlyMinimum +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  Migration +  RelativePoverty +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  Migration +  ElderlyMinimum +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Employees +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  ElderlyRetired +  Migration +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  Roma,
                   Durbeq =  ~ FamilyDoctors +  ElderlyRetired +  Migration +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  ElderlyRetired +  Migration +  RelativePoverty +  ElderlyMinimum +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  ElderlyRetired +  Migration +  RelativePoverty +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  ElderlyRetired +  Migration +  ElderlyMinimum +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  ElderlyRetired +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors +  Migration +  RelativePoverty +  ElderlyMinimum +  PrimaryEd4 +  Roma +  LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + PrimaryEd4,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + RelativePoverty + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + ElderlyMinimum + PrimaryEd4,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + ElderlyMinimum + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + ElderlyMinimum + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + Migration + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + RelativePoverty + ElderlyMinimum + PrimaryEd4,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + RelativePoverty + ElderlyMinimum + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + RelativePoverty + ElderlyMinimum + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + RelativePoverty + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + RelativePoverty + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + RelativePoverty + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + ElderlyMinimum + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + ElderlyMinimum + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + ElderlyMinimum + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyRetired + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + RelativePoverty + ElderlyMinimum + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + RelativePoverty + ElderlyMinimum + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + RelativePoverty + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + RelativePoverty + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + RelativePoverty + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + ElderlyMinimum + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + ElderlyMinimum + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + ElderlyMinimum + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + Migration + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + Employees + RelativePoverty + ElderlyMinimum + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + RelativePoverty + ElderlyMinimum + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + RelativePoverty + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Employees + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + Roma,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + RelativePoverty + ElderlyMinimum + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + RelativePoverty + ElderlyMinimum + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + RelativePoverty + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + ElderlyRetired + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma,
                   Durbeq =  ~ FamilyDoctors + Migration + RelativePoverty + ElderlyMinimum + PrimaryEd4 + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Migration + RelativePoverty + ElderlyMinimum + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Migration + RelativePoverty + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + Migration + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020,
                   Durbeq =  ~ FamilyDoctors + RelativePoverty + ElderlyMinimum + PrimaryEd4 + Roma + LivingSpace2020)

for (lagged_x in lagged_x_list) {
  SararHetDurbT1A <- spreg(eq1, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=lagged_x, model = "sarar")
  if (sign(SararHetDurbT1A$coefficients[length(SararHetDurbT1A$coefficients)])==sign(SararHetDurbT1A$coefficients[length(SararHetDurbT1A$coefficients)-1])){
    print(lagged_x)
  }
}
```

Select the durbin equations from above (which do not result in a different signs in the spatial error term and spatial lag of y) and repeat previous step to get all combinations of lagged spatial variables, for a model comparison using the AIC and BIC:
```{r}
lagged_x_list <- c(
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty + ElderlyMinimum + Roma,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum + Roma,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + ElderlyMinimum + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + Migration + ElderlyMinimum + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + Roma + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + ElderlyMinimum + Roma + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + Migration + ElderlyMinimum + Roma + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + ElderlyMinimum,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + Roma,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + ElderlyMinimum + Roma,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + ElderlyMinimum + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Roma + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + Migration + ElderlyMinimum + Roma,
  Durbeq =  ~ FamilyDoctors + Migration + ElderlyMinimum + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + Migration + Roma + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyMinimum + Roma + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration + LivingSpace2020, #best combiantion of spatially lagged x terms
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + ElderlyMinimum + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + Migration + ElderlyMinimum + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + Migration + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyMinimum + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Migration,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + ElderlyMinimum,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + Roma,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + Migration + Roma,
  Durbeq =  ~ FamilyDoctors + Migration + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyMinimum + Roma,
  Durbeq =  ~ FamilyDoctors + ElderlyMinimum + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + Roma + LivingSpace2020,
  Durbeq =  ~ FamilyDoctors + ElderlyRetired + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + Migration + ElderlyMinimum,
  Durbeq =  ~ FamilyDoctors + Migration + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + ElderlyMinimum + RelativePoverty,
  Durbeq =  ~ FamilyDoctors + Roma + RelativePoverty)

for (lagged_x in lagged_x_list) {
  SararHetDurbT1A <- spreg(eq1, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=lagged_x, model = "sarar")
  SararHetDurbT2A <- spreg(eq3, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=lagged_x, model = "sarar")
  lagged_x_restricted <- update(lagged_x, ~. -ElderlyRetired)
  SararHetDurbT1B <- spreg(eq2, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=lagged_x_restricted, model = "sarar")
  SararHetDurbT2B <- spreg(eq4, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=lagged_x_restricted, model = "sarar")
  
  #check if lambda and rho have the same sign for all models

  if (
    as.numeric(sign(SararHetDurbT1A$coefficients[length(SararHetDurbT1A$coefficients)])==sign(SararHetDurbT1A$coefficients[length(SararHetDurbT1A$coefficients)-1]))+
    as.numeric(sign(SararHetDurbT2A$coefficients[length(SararHetDurbT2A$coefficients)])==sign(SararHetDurbT2A$coefficients[length(SararHetDurbT2A$coefficients)-1]))==2)
    # as.numeric(sign(SararHetDurbT1B$coefficients[length(SararHetDurbT1B$coefficients)])==sign(SararHetDurbT1B$coefficients[length(SararHetDurbT1B$coefficients)-1])) +
    # as.numeric(sign(SararHetDurbT2B$coefficients[length(SararHetDurbT2B$coefficients)])==sign(SararHetDurbT2B$coefficients[length(SararHetDurbT2B$coefficients)-1]))==4)
    {
    print(lagged_x)
    residuals_SararHetDurbT1A <- residuals(SararHetDurbT1A) # Compute the residuals
    sige2_SararHetDurbT1A <- sum(residuals_SararHetDurbT1A^2) / length(residuals_SararHetDurbT1A) # Estimate the residual variance
    loglik_SararHetDurbT1A <- sum(dnorm(residuals_SararHetDurbT1A, mean = 0, sd = sqrt(sige2_SararHetDurbT1A))) # Compute the log-likelihood
    num_params_SararHetDurbT1A <- length(coef(SararHetDurbT1A)) # Calculate the number of parameters in the model
    n <- length(residuals_SararHetDurbT1A)
    aic_SararHetDurbT1A <- -2 * loglik_SararHetDurbT1A + 2 * num_params_SararHetDurbT1A # Compute AIC
    bic_SararHetDurbT1A <- -2 * loglik_SararHetDurbT1A + num_params_SararHetDurbT1A * log(n) # Compute BIC
    print('#MODEL 1#')
    print(paste0("AIC: ", aic_SararHetDurbT1A)) # Print AIC
    print(paste0("BIC: ", bic_SararHetDurbT1A))  # Print BIC
    
    residuals_SararHetDurbT2A <- residuals(SararHetDurbT2A) # Compute the residuals
    sige2_SararHetDurbT2A <- sum(residuals_SararHetDurbT2A^2) / length(residuals_SararHetDurbT2A) # Estimate the residual variance
    loglik_SararHetDurbT2A <- sum(dnorm(residuals_SararHetDurbT2A, mean = 0, sd = sqrt(sige2_SararHetDurbT2A))) # Compute the log-likelihood
    num_params_SararHetDurbT2A <- length(coef(SararHetDurbT2A)) # Calculate the number of parameters in the model
    n <- length(residuals_SararHetDurbT2A)
    aic_SararHetDurbT2A <- -2 * loglik_SararHetDurbT2A + 2 * num_params_SararHetDurbT2A # Compute AIC
    bic_SararHetDurbT2A <- -2 * loglik_SararHetDurbT2A + num_params_SararHetDurbT2A * log(n) # Compute BIC
    print('#MODEL 3#')
    print(paste0("AIC: ", aic_SararHetDurbT2A)) # Print AIC
    print(paste0("BIC: ", bic_SararHetDurbT2A))  # Print BIC
  }
}
```


```{r}
Durbeq =~FamilyDoctors + ElderlyRetired + Employees + Size100
Durbeq_restricted <- update(Durbeq, ~. -ElderlyRetired-Employees)
# Models-
SararHetDurbT1A <- spreg(eq1, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=Durbeq, model = "sarar")
SararHetDurbT1B <- spreg(eq2, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=Durbeq_restricted, model = "sarar")
SararHetDurbT2A <- spreg(eq3, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=Durbeq, model = "sarar")
SararHetDurbT2B <- spreg(eq4, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=Durbeq_restricted, model = "sarar")

# Restricted models other tests

Durbeq_restricted =~FamilyDoctors

eq2 = VaccinationRateT1 ~ FamilyDoctors + Migration + SocialistInv + Employees
eq4 = VaccinationRateT2 ~ FamilyDoctors + Migration + SocialistInv + Employees

SararHetDurbT1B <- spreg(eq2, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=Durbeq_restricted, model = "sarar")
SararHetDurbT2B <- spreg(eq4, data=spatdata, listw = w_queen, het = TRUE, verbose = FALSE, Durbin=Durbeq_restricted, model = "sarar")

summary(SararHetDurbT1B)
summary(SararHetDurbT2B)
```


```{r}
effects_SararHetDurbT1A <- sphet::impacts(SararHetDurbT1A, listw = w_queen, R=399)
summary(effects_SararHetDurbT1A, zstats=TRUE)
effects_SararHetDurbT1B <- sphet::impacts(SararHetDurbT1B, listw = w_queen, R=399)
summary(effects_SararHetDurbT1B, zstats=TRUE)
effects_SararHetDurbT2A <- sphet::impacts(SararHetDurbT2A, listw = w_queen, R=399)
summary(effects_SararHetDurbT2A, zstats=TRUE)
effects_SararHetDurbT2B <- sphet::impacts(SararHetDurbT2B, listw = w_queen, R=399)
summary(effects_SararHetDurbT2B, zstats=TRUE)
```


```{r}
SararHetDurbT1A$SSE/(var(spatdata$VaccinationRateT1)*(length(spatdata$VaccinationRateT1)-1))
```


```{r}
# AIC and BIC for models SararHetDurbT1A
residuals_SararHetDurbT1A <- residuals(SararHetDurbT1A) # Compute the residuals
sige2_SararHetDurbT1A <- sum(residuals_SararHetDurbT1A^2) / length(residuals_SararHetDurbT1A) # Estimate the residual variance
loglik_SararHetDurbT1A <- sum(dnorm(residuals_SararHetDurbT1A, mean = 0, sd = sqrt(sige2_SararHetDurbT1A))) # Compute the log-likelihood
num_params_SararHetDurbT1A <- length(coef(SararHetDurbT1A)) # Calculate the number of parameters in the model
n <- length(residuals_SararHetDurbT1A)
aic_SararHetDurbT1A <- -2 * loglik_SararHetDurbT1A + 2 * num_params_SararHetDurbT1A # Compute AIC
bic_SararHetDurbT1A <- -2 * loglik_SararHetDurbT1A + num_params_SararHetDurbT1A * log(n) # Compute BIC
aic_SararHetDurbT1A # Print AIC
bic_SararHetDurbT1A  # Print BIC
```


```{r}
# AIC and BIC for models SararHetDurbT1B
residuals_SararHetDurbT1B <- residuals(SararHetDurbT1B) # Compute the residuals
sige2_SararHetDurbT1B <- sum(residuals_SararHetDurbT1B^2) / length(residuals_SararHetDurbT1B) # Estimate the residual variance
loglik_SararHetDurbT1B <- sum(dnorm(residuals_SararHetDurbT1B, mean = 0, sd = sqrt(sige2_SararHetDurbT1B))) # Compute the log-likelihood
num_params_SararHetDurbT1B <- length(coef(SararHetDurbT1B)) # Calculate the number of parameters in the model
n <- length(residuals_SararHetDurbT1B)
aic_SararHetDurbT1B <- -2 * loglik_SararHetDurbT1B + 2 * num_params_SararHetDurbT1B # Compute AIC
bic_SararHetDurbT1B <- -2 * loglik_SararHetDurbT1B + num_params_SararHetDurbT1B * log(n) # Compute BIC
aic_SararHetDurbT1B # Print AIC
bic_SararHetDurbT1B  # Print BIC
```

```{r}
# AIC and BIC for models SararHetDurbT2A
residuals_SararHetDurbT2A <- residuals(SararHetDurbT2A) # Compute the residuals
sige2_SararHetDurbT2A <- sum(residuals_SararHetDurbT2A^2) / length(residuals_SararHetDurbT2A) # Estimate the residual variance
loglik_SararHetDurbT2A <- sum(dnorm(residuals_SararHetDurbT2A, mean = 0, sd = sqrt(sige2_SararHetDurbT2A))) # Compute the log-likelihood
num_params_SararHetDurbT2A <- length(coef(SararHetDurbT2A)) # Calculate the number of parameters in the model
n <- length(residuals_SararHetDurbT2A)
aic_SararHetDurbT2A <- -2 * loglik_SararHetDurbT2A + 2 * num_params_SararHetDurbT2A # Compute AIC
bic_SararHetDurbT2A <- -2 * loglik_SararHetDurbT2A + num_params_SararHetDurbT2A * log(n) # Compute BIC
aic_SararHetDurbT2A # Print AIC
bic_SararHetDurbT2A  # Print BIC
```

```{r}
# AIC and BIC for models SararHetDurbT2B
residuals_SararHetDurbT2B <- residuals(SararHetDurbT2B) # Compute the residuals
sige2_SararHetDurbT2B <- sum(residuals_SararHetDurbT2B^2) / length(residuals_SararHetDurbT2B) # Estimate the residual variance
loglik_SararHetDurbT2B <- sum(dnorm(residuals_SararHetDurbT2B, mean = 0, sd = sqrt(sige2_SararHetDurbT2B))) # Compute the log-likelihood
num_params_SararHetDurbT2B <- length(coef(SararHetDurbT2B)) # Calculate the number of parameters in the model
n <- length(residuals_SararHetDurbT2B)
aic_SararHetDurbT2B <- -2 * loglik_SararHetDurbT2B + 2 * num_params_SararHetDurbT2B # Compute AIC
bic_SararHetDurbT2B <- -2 * loglik_SararHetDurbT2B + num_params_SararHetDurbT2B * log(n) # Compute BIC
aic_SararHetDurbT2B # Print AIC
bic_SararHetDurbT2B  # Print BIC
```